\Large \bf SIRIUS: A Revolutionary Approach to I/O and Storage for Exascale Scientific Computing

### 2016-03-07 papers/2016-HotStorage/paper.tex
added table for reading options

The first generation of standardized APIs for scientific computing IO focused
on serial output from a single process to a distributed or local file system.
These systems still exist in the form of serial HDF5~\cite{chilan:2006:hdf5-performance} and the
end-of-life NetCDF 3.x~\cite{netcdf3}. The key feature of these APIs was the
logical data model that eliminated any parallel data decomoposition artifacts
from the actual data storage. This offered consistent performance and regular
data access patterns.

@@

The third generation is a reaction to the second based on recognizing that
maintaining the logical data model in physical form became inordinately
expensive and offered a performance penalty in most cases when running at
scale. Libraries like ADIOS~\cite{lofstead:2009:adaptable} shifted from writing complete
variables from distributed processes as a single entity to treating all data
from a single source to be treated as a unit. This eliminated the data
reorganization at the cost of more complicated reading patterns. In spite of
this extra complication, the performance is generally
better~\cite{lofstead:2011:six-degrees}.

SIRIUS represents the next generation. New platforms are introducting
additional storage layers to accelerate IO operations and to compensate for
plateauing storage array performance. At the same time, data sizes continue
to grow. The fourth generation IO API will have to both address the IO API
and middleware layer as well as the deeper, more complex storage hierarchy.

Systems like ADIOS showed the performance advantage for decomposed data
layouts, but still rely on a uniform storage layer. SIRIUS represents a shift
to consider the local portion of each globally distributed variable as a
management entity with data placement decisions across the entire storage
hierarhy rather than on a single tier.  Decisions on what data to place where
is driven by the data contents itself according to end-user specified
priorities and system characteristics. For example, if temperature and pressure
are nearly always retrieved together, organizing the data to keep just these
two arrays together without any interleaving from other variables or scalars
will accelerate later reading. If these variables could also be ``processed''
such that the important data features are maintained in high resolution while
less interesting regions are stored in highly compressed--even using lossy
compression--forms, the amount of data written can be greatly increased. While
accelerating writing is important. This approach also supports accelerating
reading.

@@

\noindent\textbf{Open and Close} need to add additional information to include
intent information. The intent hints to the system that....

\noindent\textbf{Refactoring} may be a uniform operation across an
entire variable. For example, splitting a double into three parts--two byes,
four bytes, and two bytes, yields a two-byte value that is the majority of the
data precision. The next four bytes would be most of the rest of the data
precision. The last two bytes are the remainder of the precision. For quick
overviews, just reading the two-byte portion is sufficient to get a rough idea
of data contents. If particular data looks interesting, more and more data can
be read to improve the precision. This can easily be applied uniformly by each
process without coordinating. A more dynamic approach would be to apply a
wavelet compression~\cite{klappenecker:1995:wavelet} approach that generates different compressed
data sizes at a fixed error rate depending on the data entropy. This is still a
uniform operator, but the resulting data can be of radically different sizes.

A more complex, but interesting approach is to offer an {\em Auditor} that
performs the same computation, but with some simplification that greatly
reduces computation intensity with only a minor loss of precision over the
short term. Longer term, this auditor would need to be reset to bring it back
into alignment with the full precision simulation. The much smaller reduced
precision data could be used instead of the full precision in many
circumstances.

A middle ground approach between these two refactoring approaches is to use a
non-uniform operator across a variable. Instead of, for example, splitting on
byte boundaries to make multiple values, a function that identifies data
features through a local analytics function could identify ``interesting'' data
that should be maintained in full precision while other areas can be more
highly compressed. Such an operator would decompose the data from a process
further than a single process or node depending on the data features. This
approach may balance precision better than wavelet compression, but at a higher
analytics cost.

We plan to investigate all three approaches as different applications will have
different techniques that work best. Any performant storage system should
offer the flexibility to incorporate application specific data refactoring
techniques and sufficient metadata support for an arbitrary client to access
the data successfully.

\noindent\textbf{Quality of Service} deadlines are important for driving how
much time is spent performing IO. At the same time, data quality must be
maintained at a sufficient level for the resulting science to be possible.
The system must offer a way for an end-user to negotiate with the system for
trading off between time spent in IO and the amount of data written or read.
Part of that tradeoff will be increased or reduced data quality, depending on
the system.

>>>>

High performance computing simulations have driven tightly coupled platform
scaling for decades. The output from these simulations are analyzed using
regular tools and techniques optimized for this environment.  The first
generation of standardized APIs for scientific computing IO focused on serial
output from a single process to a distributed or local file system.  These
systems still exist in the form of serial
HDF5~\cite{chilan:2006:hdf5-performance} and the end-of-life NetCDF
3.x~\cite{netcdf3}. The key feature of these APIs was the logical data model
that eliminated any parallel data decomoposition artifacts from the actual data
storage. This offered consistent performance and regular data access patterns.

@@

The third generation is a reaction to the second based on recognizing that
maintaining the logical data model in physical form became inordinately
expensive~\cite{lofstead:2011:six-degrees} and offered a performance penalty in
most cases when running at scale. Libraries like
ADIOS~\cite{lofstead:2009:adaptable} shifted from writing complete variables
from distributed processes as a single entity to treating all data from a
single source to be treated as a unit. This eliminated the data reorganization
at the cost of more complicated reading patterns. In spite of this extra
complication, the performance is generally
better~\cite{lofstead:2011:six-degrees}.

SIRIUS represents the next generation. New platforms are introducting
additional storage layers to accelerate IO operations and to compensate for
plateauing storage array performance. At the same time, data sizes continue
to grow. The fourth generation IO API will have to both address the IO API
and middleware layer as well as the deeper, more complex storage hierarchy.
SIRIUS considers the local portion of each globally distributed variable as a
management entity with data placement decisions across the entire storage
hierarhy rather than on a single tier.  Decisions on what data to place where
is driven by the data contents itself according to end-user specified
priorities and system characteristics. For example, if temperature and pressure
are nearly always retrieved together, organizing the data to keep just these
two arrays together without any interleaving from other variables or scalars
will accelerate later reading. If these variables could also be ``processed''
such that the important data features are maintained in high resolution while
less interesting regions are stored in highly compressed--even using lossy
compression--forms, the amount of data written can be greatly increased. While
accelerating writing is important. This approach also supports accelerating
reading.

@@

\noindent\textbf{Refactoring} may be a uniform operation across an
entire variable. For example, splitting a double into three parts--two byes,
four bytes, and two bytes, yields a two-byte value that is the majority of the
data precision. The next four bytes would be most of the rest of the data
precision. The last two bytes are the remainder of the precision. For quick
overviews, just reading the two-byte portion is sufficient to get a rough idea
of data contents. If particular data looks interesting, more and more data can
be read to improve the precision. This can easily be applied uniformly by each
process without coordinating. From a storage hierarchy perspective, the primary
value portion could be written to high performance, low capacity storage while
the rest is written to slower storage. That would reduce the data quantity
written to NVM, for example, by 75\% for a 2/6 split enabling far more output
to be written to this fast tier with the rest spooled asynchronously.

A more complex, but interesting approach is to offer an {\em Auditor} that
performs a similar, but lower precision computation that greatly reduces
computation intensity with only a minor loss of precision over short spatial
and temporal domains. As the simulation progresses, the auditor drift would
need to be reset to bring it back into alignment. The much smaller, reduced
precision data could be used instead of the full precision in many
circumstances.  A more dynamic lossy approach would be to apply a wavelet
compression~\cite{klappenecker:1995:wavelet} approach that generates different
compressed data sizes at a fixed error rate depending on the data entropy.
These are still uniform operators, but the resulting data can be of radically
different sizes. For this approach, some or all of the lower precision data
can be written to the fast tier while full precision can be spooled
asynchrnously to the slower tier(s).

A middle ground approach between these two refactoring approaches is to use a
non-uniform operator across a variable. Instead of, for example, splitting on
byte boundaries to make multiple values, a function that identifies data
features through a local analytics function could identify ``interesting'' data
that should be maintained in full precision while other areas can be more
highly compressed. Such an operator would decompose the data from a process
further than a single process or node depending on the data features. This
approach may balance precision better than wavelet compression, but at a higher
analytics cost. We would put data areas with interesting features on the fast
tier while the rest could be spooled to the slow tier asynchronously.

We plan to investigate all three approaches as different applications will have
different techniques that work best. Our initial tests have demonstrated the
potential for the auditor approach in an isolated environment. We have also
built a prototype ADIOS transport method for the byte-split apporach. Any
performant storage system should offer the flexibility to incorporate
application specific data refactoring techniques and sufficient metadata
support for an arbitrary client to access the data successfully.

\noindent\textbf{Quality of Service} deadlines are important for driving how
much time is spent performing IO. At the same time, data quality must be
maintained at a sufficient level for the resulting science to be possible.
The system must offer a way for an end-user to negotiate with the system for
trading off between time spent in IO and the amount of data written or read.
Part of that tradeoff will be increased or reduced data quality, depending on
the system. For example, when data is written, a few different forms may be
output given sufficient space and time. One may be every fourth element
across a regular mesh reduces the data size to $1/64^th$ the full data set.
Further dividing that using byte-splitting can reduce it by another 75\%. For
quick overviews to search data sets, this reduced form can be sufficient
driving which data sets should be investigated in more detail. The key driving
factor is time. The applicaiton scientist can read 0.4\% of the full data
volume with a similar time reduction. SIRIUS will offer a negotiation process
to read any data.

First, a user will request a particular variable and timesetp. SIRIUS will
return a table similar to Table~\ref{tab:results}. For a given variable ``A'',
the end-user can then select the data quality and approximate reading time for
the operation. In this case, there are 5 options for obtaining the whole data
set (1/2/3/4, 5/6, 7/8, 9, 10). The first four represent a spatial and byte
split, 5/6 represent just strided writing, 7/8 represent just a byte-split, 9
represents a wavelet compressed version, and 10 is the full precision data.
The user would specify to read a particular data instance that would proceed
normally.  Quality of Service soft guarantees specified as part of the request
will result in SIRIUS allocating soft bandwidth reservations to meet the
request.

\begin{table}[tbp]
\centering
\caption{Reading Options for Variable ``A''}
\label{tab:results}
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\# & Size & Time & Time Err & Refactoring & Data Err\\
\hline
1 & $(\frac{1}{4})^3(\frac{3}{8})$A & 10s & $\pm$ 3s & stride, byte-split & 99\% \\
\hline
2 & $(\frac{3}{4})^3(\frac{3}{8})$A & 90s & $\pm$ 30s & stride, byte-split & 58\% \\
\hline
3 & $(\frac{1}{4})^3(\frac{5}{8})$A & 16s & $\pm$ 5s & stride, byte-split & 0.01\% \\
\hline
4 & $(\frac{3}{4})^3(\frac{5}{8})$A & 120s & $\pm$ 50s & stride, byte-split & 0.01\% \\
\hline
\hline
5 & $(\frac{1}{4})^3$A & 1200s & $\pm$ 30s & stride & 98\% \\
\hline
6 & $(\frac{3}{4})^3$A & 2400s & $\pm$ 90s & stride & 58\% \\
\hline
\hline
7 & $(\frac{3}{8})$A & 1350s & $\pm$ 120s & byte-split & 5\% \\
\hline
8 & $(\frac{5}{8})$A & 2250s & $\pm$ 120s & byte-split & 0.01\% \\
\hline
\hline
9 & A & 36s & $\pm$ 6s & wavelet & 1\% \\
\hline
\hline
10 & A & 3600s & $\pm$ 600s & none & 0\% \\
\hline
\end{tabular}
\end{table}
